---
title: "**Data Visualization with R ggplot2**"
author: "Andriy Gazin"
date: "August 11, 2016"
output: 
  slidy_presentation:
    theme: cerulean
---
#**Load the data**

```{r eval=TRUE, echo=TRUE, message=FALSE}
Sys.setlocale("LC_TIME", "en_US.UTF-8")
weather <- read.csv("https://raw.githubusercontent.com/andriy-gazin/datasets/master/kyiv_weather_2015.csv",
                    stringsAsFactors = FALSE)
str(weather)
```

#**Inspect summary of the data**

```{r eval=TRUE, echo=TRUE, message=FALSE}
summary(weather)
```

#**Add new variables**

```{r, message=FALSE}
library(dplyr)
library(lubridate)
weather <- weather %>% 
  mutate(date = as.POSIXct(date, tz = "UTC+2"),
         month = month(date, label = TRUE, abbr = TRUE))

for (i in 1:nrow(weather)) {
if(weather$month[i] %in% c("Dec", "Jan", "Feb")) {weather$season[i] = "Winter"}
else if(weather$month[i] %in% c("Mar", "Apr", "May")) {weather$season[i] = "Spring"}
else if(weather$month[i] %in% c("Jun", "Jul", "Aug")) {weather$season[i] = "Summer"}
else if(weather$month[i] %in% c("Sep", "Oct", "Nov")) {weather$season[i] = "Autumn"}
}
```

#**ggplot2 compared to base**

```{r, eval=TRUE, echo=TRUE, message=FALSE}
plot(x = weather$temp, y = weather$hum)

```

```{r, eval=TRUE, echo=TRUE, message=FALSE}
library(ggplot2)
ggplot(weather)+geom_point(aes(x = temp, y = hum))

```

#**Layered grammar of graphics**

**Every plot consists of:**

* Data  
* Aesthetic mappings  
* Geometry  
* Statistics  
* Scales  
* Theme  

#**ggplot object**

```{r, eval=FALSE}
ggplot(weather, aes(x = temp))

```

```{r, eval=TRUE}
g <- ggplot(weather, aes(x = temp))

```

#**What's in the box?**

```{r}
str(g)
```

#**Aesthetics**

```{r, eval=TRUE, echo=TRUE, message=FALSE, cache=TRUE}
ggplot(weather, aes(x = temp))

```

#**All aesthetics**

```{r}
ggplot2:::.all_aesthetics
```

#**Geometry**

```{r}
ls(pattern = '^geom_', env = as.environment('package:ggplot2'))

```

#**Data + Aesthetics + Geometry**

```{r, eval=TRUE, echo=TRUE, message=FALSE}
ggplot(weather, aes(x = temp))+
  geom_histogram()+
  geom_rug()
```

#**Histogram - Adjust number of bins**
```{r}
ggplot(weather, aes(x = temp))+
  geom_histogram(bins = max(weather$temp) - min(weather$temp))
```

#**Histogram - Adjust binwidth**
```{r}
ggplot(weather, aes(x = temp))+
  geom_histogram(binwidth = 1, fill = "brown", alpha = 0.75)
```

#**Histogram - Adjust binwidth & color by season**
```{r}
ggplot(weather, aes(x = temp, fill = season))+
  geom_histogram(binwidth = 1, alpha = 0.75)
```

#**Histogram - Add scales**
```{r}
ggplot(weather, aes(x = temp, fill = season))+
  geom_histogram(binwidth = 1, alpha = 0.75)+
  scale_x_continuous(breaks = seq(min(weather$temp), max(weather$temp), 10),
                     minor_breaks = FALSE)+
  scale_fill_brewer(type = "qual", palette = 2)
```

#**Histogram - Change theme**
```{r}
ggplot(weather, aes(x = temp, fill = season))+
  geom_histogram(binwidth = 1, alpha = 0.75)+
  scale_x_continuous(breaks = seq(min(weather$temp), max(weather$temp), 10),
                     minor_breaks = FALSE)+
  scale_fill_brewer(type = "qual", palette = 2)+
  theme_minimal()
```

#**Histogram - Change geometry**
```{r}
ggplot(weather, aes(x = temp, color = season))+
  geom_freqpoly(binwidth = 1, alpha = 0.75, size = 1)+
  scale_x_continuous(breaks = seq(min(weather$temp), max(weather$temp), 10),
                     minor_breaks = FALSE)+
  scale_color_brewer(type = "qual", palette = 2)+
  theme_minimal()
```

#**Histogram - Change geometry once again**
```{r}
ggplot(weather, aes(x = temp, color = season))+
  geom_density(alpha = 0.75, size = 1)+
  scale_x_continuous(breaks = seq(min(weather$temp), max(weather$temp), 10),
                     minor_breaks = FALSE)+
  scale_color_brewer(type = "qual", palette = 2)+
  theme_minimal()
```

#**Histogram - Facet**
```{r}
ggplot(weather, aes(x = temp))+
  geom_histogram(binwidth = 1, alpha = 0.75)+
  facet_wrap(~season, nrow = 2, ncol = 2)+
  theme_minimal()
```

#**Boxplot**
```{r}
ggplot(weather)+
  geom_boxplot(aes(x = 1, y = temp))
```

#**Boxplot - Add jitter points**
```{r}
ggplot(weather, aes(x = 1, y = temp))+
  geom_jitter(alpha = 0.1, color = 'gray50')+
  geom_boxplot(fill = NA)
  
```

#**Boxplot - Adjust width**
```{r}
ggplot(weather, aes(x = 1, y = temp))+
  geom_jitter(alpha = 0.05, color = 'gray40', width = 0.2, height = 0.1)+
  geom_boxplot(width = 0.2, fill = NA, size = 0.75, color = 'darkred')
  
```

#**Boxplot - Spit by month**
```{r}
ggplot(weather, aes(x = month, y = temp))+
  geom_jitter(alpha = 0.05, color = 'gray40', height = 0.1)+
  geom_boxplot(fill = NA, color = 'darkred')
  
```

#**Boxplot - Spit by season**
```{r}
ggplot(weather, aes(x = season, y = temp))+
  geom_jitter(alpha = 0.05, color = 'gray40', height = 0.1)+
  geom_boxplot(fill = NA, color = '#b02350')
  
```

#**Boxplot - Spit by season correctrly**
```{r}
weather$season <- factor(weather$season, levels = c('Winter', 'Spring', 'Summer', 'Autumn'),
                         ordered = TRUE)

ggplot(weather, aes(x = season, y = temp))+
  geom_jitter(alpha = 0.05, color = 'gray40', height = 0.1)+
  geom_boxplot(fill = NA, color = '#b02350')
  
```

#**Boxplot - Change geometry**
```{r}
ggplot(weather, aes(x = season, y = temp))+
  geom_violin()
  
```

#**Scatterplot**
```{r}
ggplot(weather, aes(temp, hum))+
  geom_point()
```

#**Scatterplot - Color by season**
```{r}
ggplot(weather, aes(temp, hum, color = season), alpha = 0.5)+
  geom_point()
```

#**Scatterplot - Change colorscale**
```{r}
ggplot(weather, aes(temp, hum, color = season))+
  geom_jitter(alpha = 0.5)+
  scale_color_brewer(type = 'qual', palette = 2)
```

#**Scatterplot - Add regression line**
```{r}
ggplot(weather, aes(temp, hum, color = season))+
  geom_jitter(alpha = 0.5)+
  geom_smooth(method = "lm")+
  scale_color_brewer(type = 'qual', palette = 2)
```

#**Scatterplot - Add regression line**
```{r}
ggplot(weather, aes(temp, hum))+
  geom_jitter(aes(color = season), alpha = 0.5)+
  geom_smooth(method = "lm")+
  scale_color_brewer(type = 'qual', palette = 2)
```

#**Scatterplot - Facet**
```{r}
ggplot(weather, aes(temp, hum))+
  geom_jitter(alpha = 0.25, color = 'gray50')+
  geom_smooth(method = "lm", se = FALSE, color = 'brown')+
  facet_wrap(~season)
```

#**Scatterplot - Change geometry**
```{r}
ggplot(weather, aes(temp, hum))+
  geom_count(alpha = 0.25, color = 'gray50')+
  geom_smooth(method = "lm", se = FALSE, color = 'brown')+
  facet_wrap(~season)
```

#**Scatterplot - Change geometry**
```{r}
ggplot(weather, aes(temp, hum))+
  geom_density2d(alpha = 0.50)+
  facet_wrap(~season)
```

#**Scatterplot - Change geometry**
```{r}
ggplot(weather, aes(temp, hum))+
  geom_hex(alpha = 0.50)+
  viridis::scale_fill_viridis()+
  facet_wrap(~season)
```

#**Line chart**
```{r}
ggplot(weather, aes(date, temp))+
  geom_line()
```

#**Line chart - Facet**
```{r}
ggplot(weather, aes(date, temp))+
  geom_line()+
  facet_wrap(~month, scales = "free_x", nrow = 4, ncol = 3)
```

#**Line chart - Aggregate data**
```{r}
weatherAgr <- weather %>% 
  mutate(date = as.Date(date)) %>% 
  group_by(date) %>% 
  summarise(min = min(temp),
            max = max(temp),
            temp = mean(temp))

ggplot(weatherAgr, aes(date, temp))+
  geom_line()
```

#**Line chart - Add scales**
```{r}
ggplot(weatherAgr, aes(date, temp, color = temp))+
  geom_line()+
  viridis::scale_color_viridis()+
  scale_x_date(date_breaks = "1 month", date_labels = "%B")
```

#**Line chart - Change geometry**
```{r}
ggplot(weatherAgr)+
  geom_linerange(aes(x = date, ymin = min, ymax = max, color = temp))+
  viridis::scale_color_viridis()+
  scale_x_date(date_breaks = "1 month", date_labels = "%B")+
  theme_minimal()
```

#**Line chart - Change coordinate system**
```{r}
ggplot(weatherAgr)+
  geom_linerange(aes(x = date, ymin = min, ymax = max, color = temp))+
  viridis::scale_color_viridis()+
  scale_x_date(date_breaks = "1 month", date_labels = "%B")+
  coord_polar()+
  theme_minimal()
```

#**Bar chart**
```{r}
conditions <- weather %>% 
  group_by(cond) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  filter(count > 100)

ggplot(conditions, aes(cond, count))+
  geom_bar(stat = "identity")
```

#**Bar chart - Sort & make horizontal**
```{r}
ggplot(conditions, aes(reorder(cond, count), count))+
  geom_bar(stat = "identity", fill = 'brown')+
  coord_flip()+
  ggthemes::theme_fivethirtyeight()
```
